Geometric variational problems

### 2020-11-11 NSF2020/Description/background.tex
rjr

\begin{wrapfigure}[16]{r}{0.22\textwidth}
\centerline{\includegraphics[width=0.22\textwidth]{figures/BG_fig1.pdf}}
\caption{ \footnotesize
A collection $P_1,$ $P_2,$ $P_3,$ \dots of rigid particles. The exterior domain
$\Omega$ represents the solvent.}
  \label{fig:domain}
\end{wrapfigure}
Recently PIs RR and YNY developed a mathematical model called the
hydrophobic attraction potential (HAP) \cite{Fu2018_SIAM} that is based
on the physical origin of hydrophobicity.  This model addresses the
major shortcomings of molecular dynamics (MD) and continuum approaches.
Based on preliminary results (\S\ref{sec:preliminary_work}) the PIs
propose to extend this HAP model to offer an alternative modeling
methodology that leads to new mathematical ideas, and is both physically
accurate and computationally practical.
%
%
%The word hydrophobic (water fearing) derives from the low solubility of oil (hydrocarbon solute) in water and vice versa. 
%It causes hydrophobic moieties to aggregate and cluster,
%is responsible for the adhesion between hydrophobic surfaces \cite{Ducker2016}, large contact angles on a 
%dewetting surface \cite{Arenas2019,Sandre1999}, accumulation of particles along interfaces \cite{Lee2013,Lee2014}, 
%formation of micelles and bilayers \cite{Israelachvili80}, and protein folding and membrane insertion \cite{Kabelka2018}.
%
%Lipids are amphiphilic molecules whose  structure possesses
%both hydrophobic and hydrophilic parts. 
%The amphiphilic property is what allows lipids to form the membranes and 
%compartments of living cells \cite{Israelachvili80}. 
%More specifically, a lipid consists of 
%an elongated hydrocarbon tail that is hydrophobic, attached to a polar head that is hydrophilic.
%To shield the hydrophobic tails from water, lipids self-assemble into micelles and bilayers. 
%A micelle is a spherical arrangement of lipids with tails terminating at the micelle center. 
%A bilayer consists of two layers of lipids called monolayers, where the lipid tails point 
%from the monolayer surface into the bilayer core. 
%
%The mathematical modeling of a biological membrane is a challenging problem in applied mathematics. 
%Bilayers are elastic and resist deformations like bending, twisting, and stretching.
%Their elastic deformations are well described by the theory of liquid crystals \cite{ANDRIENKO2018520}.
%Lipid bilayer membranes can also be fluidic, and the lateral translation of lipids (or any membrane bound
%proteins) couples nonlocaly to the motion of the aqueous environment \cite{MerkelSackmannEvans1989,StoneAjdari1998_JFM,OppenheimerDiamant2009_BJ,OppenheimerDiamant2011_PRL}. 
%%Finally, the membranes of cellular 
%%compartments are constantly merging and pinching as part of intracellular trafficking. 
%%Therefore monolayer surfaces undergo discontinuous deformations. 
%
%There are two prevailing approaches in membrane modeling, and each has its advantages
%and disadvantages. Molecular dynamics (MD) is to date the only tool capable of resolving granular biological details
%The computational cost of MD, however, grows with the sixth power of the sample diameter and 
%so simulations are severely limited to 
%small system sizes and short time scales \cite{DiCarlo2019}.
%The other approach, continuum mechanics, assumes smooth surfaces and can therefore model 
%realistic systems over physical times. 
%But continuum description of a biological membrane ignores the granularity of lipid molecules, and thus requires some assumptions when a
%membrane ruptures or when two membranes fuse \cite{ChKo08}. 
%%
%%mechanics presumes, rather than predicts, the nucleation of discontinuities. 
%%
%
he proposed research aims to provide fundamental understanding of the self-assembly dynamics of amphiphilic particles. These results will facilitate optimal design of smart materials by
tuning the geometry and properties of the amphiphilic particles.

\subsection{Hydrophobic Attraction Potential (HAP)}
\label{sec:HAP}
%Based on the physical origin of hydrophobicity,  we have devised a
%functional, called the hydrophobic attraction potential (HAP), to model
%hydrophobic forces~\cite{Fu2018_SIAM}. 
The motivation for the HAP concept stemmed from PI RR's earlier work on
the problem of energy barriers in membrane
fusion~\cite{RyKlYaCo16,Chetal16}. By applying a squared gradient
mathematical theory for hydrophobic attraction between planar
surfaces~\cite{Eriksson1989,Lum1999,Menshikov2017,Marcelja1977}, PI RR
and collaborators resolved the long-standing issue of accounting for the
energy of a monolayer fissure surface during topological transitions.
Based on PI RR and YNY's coarse-grained membrane modeling
work~\cite{Fu2017}, the investigators initiated a collaboration to
create a gradient theory for arbitrary collections of hydrophobic and
amphiphilic particles. This new method eliminates the costly calculation
of water by treating the solvent implicitly, and avoids complicated
re-meshing schemes of continuum approaches by utilizing a particle-based
representation.

To define HAP, consider a region $\Omega \subset \mathbb{R}^3$  modeling
the solvent surrounding a number of particles (Figure \ref{fig:domain}).
The HAP must be an energy of the shape of the region because hydrophobic
attraction is non-additive. The boundary of the region is the
water-particle interface and some parts of this surface are hydrophobic
while others are hydrophilic. Monte Carlo simulations show that water
changes structure at hydrophobic interfaces~\cite{Luzar1987,
Jonsson2006, Varilly2011}. Due to rapid fluctuation in the hydrogen bond
network, restructuring at the interface extends into bulk water. This
motivates the following definition for the HAP of the particles:
\begin{align}
\label{HAP}
  \Phi = \gamma \iiint_{\Omega} \rho |\nabla u|^2 + \rho^{-1}u^2 \,dx \,dy \,dz, 
\end{align}
where $u$ is the unique solution of the screened Laplace equation boundary value problem 
\begin{equation}
  \label{SL}
  -\rho^2 \Delta u + u = 0, \mbox{ } \mathbf{x} \in \Omega, \qquad
  u = f,  \mbox{ } \mathbf{x} \in \partial \Omega, \qquad 
  u(\mathbf{x}) \to 0 \mbox{ as } |\mathbf{x}| \to \infty.
\end{equation}
The scalar function $u(\mathbf{x})$ is called the water activity. Its
boundary values $f$ define the degree of hydrophobicity of the
water-particle interface. If $f=1$, the interface is hydrophobic and
water has lost rotational freedom. Conversely, if $f=0$, the interface
has the same rotational freedom as in bulk water. The parameter $\rho$
is the decay length of attraction, around 1 nm, and the parameter
$\gamma$ is the interfacial tension~\cite{Israelachvili1954}. Both
parameters contain information about the quality of the
solvent~\cite{Discher2002}.

From a phenomenological perspective, equations~\eqref{HAP}
and~\eqref{SL} form an appropriate model of hydrophobic attraction.
Specifically, using boundary layer analysis, it is possible to show that
$\Phi$ converges to a surface energy in the zero-decay length
limit~\cite{Lee2018, Lin2015, Shibata2004}. For non-zero decay lengths,
solutions of~\eqref{SL} yield an attractive interaction between
hydrophobic parallel disks separated by water~\cite{Eriksson1989}. Also,
we have demonstrated with simulation that the forces derived from HAP
theory are non-additive and significantly deviate from a pairwise
potential~\cite{Meyer2006, Fu2018_SIAM}. 
%The hydrophobic force has been implicated in the directed folding of proteins, 
%adhesion between biological membranes, but it is still unanswered as to whether
%the hydrophobic force of the form (\ref{HAP}--\ref{SL}) exists between small molecules.

@@

To define the particle dynamics, consider the exterior domain $\Omega$
for a finite number of disjoint, rigid and closed particles $P_1$,
$P_2$, \ldots, $P_n$, each with Lipschitz boundary, and boundary data
$f$ in the Sobolev space $H^1(\Omega)$. For any fixed configuration, the
minimizer of the functional~\eqref{HAP} among all $u - f$ with vanishing
trace is the solution of~\eqref{SL}. Conversely, from maximum principles
and energy estimates, the solution of~\eqref{SL} is the minimizer
of~\eqref{HAP}, and this supplies $\Phi$ with a well-defined value.
Taking the derivative of~\eqref{HAP} with respect to particle
configurations, and using the boundary value problem~\eqref{SL}, the
first variation of the hydrophobic potential $\Phi$ yields a symmetric,
rank-two tensor~\cite{Bandle2015, Schiffer1954, Grinfeld2010} called the
hydrophobic stress:

@@

  \mathbf{F}_i = \iint_{\partial P_i} \boldsymbol{\sigma}_{\text{hydro}}
  \cdot \boldsymbol{\nu} \,dS,\quad
  \mathbf{T}_i = \iint_{\partial P_i} \mathbf{x} \times
  (\boldsymbol{\sigma}_{\text{hydro}} \cdot \boldsymbol{\nu}) \,dS.

@@

Our previous work verifies that this system is force- and torque-free
and \S\ref{subsec:specific_aim_2} addresses the numerical challenges in
evaluating~\eqref{stress} and how these challenges are overcome. To
supply viscous dissipation, we incorporate the mobility problem in
Stokes flow to obtain the rigid body motions:

@@

\label{eq:stokes}
\begin{aligned}
  &-\mu \Delta \mathbf{u} + \nabla p = 0, \quad \mathbf{x} \in \Omega, \qquad 
  \nabla \cdot \mathbf{u} = 0,  \quad \mathbf{x} \in \Omega,\\
  &{\bf u}(\mathbf{x}) \to 0 \quad \text{as}\ |\mathbf{x}|\to \infty,\qquad 
  \mathbf{u}(\mathbf{x})|_{\partial P_i} = \mathbf{v}_i +
\boldsymbol{\omega}_i\times(\mathbf{x} - \mathbf{a}_i),\\
&\int_{\partial P_i}\boldsymbol{\sigma}\cdot {\bf n} dS=-{\bf F}_i, \quad
\int_{\partial P_i}(\mathbf{x}-\mathbf{a}_i)\times (\boldsymbol{\sigma} \cdot \mathbf{n}) dS=-{\bf T}_i.
\end{aligned}

@@

Here $\mu$ is the fluid viscosity and the first two equations state that
the fluid motion is a divergence-free Stokes flow; the third equation
specifies that the fluid velocity vanishes at infinity; the fourth
equation enforces a rigid body motion on each particle, where
$\mathbf{v}_i$ and $\boldsymbol{\omega}_i$ are unknown translation and
angular velocities of the $i$-th particle and $\mathbf{a}_i$ is the
center of mass of $P_i$; the last two equations state that the net
forces and torques on each particle are $\mathbf{F}_i$ and
$\mathbf{T}_i$ from~\eqref{forceandtorque}, respectively, and
$\boldsymbol{\sigma}$ is the fluid shear stress. The time integration of
particle configurations goes as follows: \textbf{(i)} solve the
BVP~\eqref{SL} for the screened Laplace equation, \textbf{(ii)}
determine the rigid body forces and torques~\eqref{forceandtorque},
\textbf{(iii)} solve the Stokes mobility problem~\eqref{eq:stokes} for
rigid body motion, \textbf{(iv)} update the particle configuration.
Numerical challenges and how these challenges are overcome are described
in \S\ref{subsec:specific_aim_2}.

@@

Our simulations use Janus particles to model lipid amphiphiles which are
popular in material science and physics for creating functional
materials~\cite{Lee2014, Lee2013}. Janus particles are typically
spherical with a biphasic material label on either hemisphere, endowing
the particle with a directional order. We considered a two-dimensional
system, and represented an elongated lipid by elliptical particles with
hydrophobic and hydrophilic labels along the ellipse's long axis. Under
the hydrophobic force, and with an excluded volume to prevent
collisions, the Janus particles spontaneously merge and realign to form
bilayers. This occurs only as a result of energy minimization and does
not require artificial inputs. To our knowledge, this is the first
demonstration of bilayer self-assembly by a continuum hydrophobic
interaction model~\cite{Noguchi2001, Farago2003, Brannigan2006,
Brooks2009, Wang2013}.

It is worth emphasizing that the HAP model uses only a few parameters;
interfacial tension, decay length, repulsion strength, and particle
shape. For example, an elastic modulus for stretching a vesicle from
micropipette manipulation directly calibrate our interfacial tension
parameter. This is in direct contrast with pair-potential based
approaches in MD simulations and coarse-grained models where many more
parameters are required~\cite{Varilly2011, Wang2013}.
%MD simulators have also made measurements and lately there is better and better agreement with reality. 
%But even the simplest coarse grained models based on pair potentials for lipids has many more parameters \cite{Varilly2011,Wang2013} .

As a proof of concept, our work has already tested for elastic energies
for bending, stretching, and tilt of the bilayer assembly. Extracting
coefficients from the HAP simulations shows strikingly positive
agreement with experimentally determined values~\cite{Fu2018_SIAM}.
Encouraged by these results and the hydrodynamic simulations of the
bilayer assembly, the PIs propose to extend the HAP model and the
numerical schemes (see \S\ref{sec:proposed-work}) to make direct
comparison with experiments, the continuum liquid crystal results and MD
simulations results, to establish that the HAP model has the capability
and adaptivity to model phenomena across length scales and time scales.

Over the past decades, researchers have used a number of mathematical
tools to simulate vesicles in a shear flow, including lattice
Boltzmann~\cite{KaouiHartingMisbah2011_PRE}, coarse-grained Brownian
dynamics~\cite{NoguchiTakasu2002_BJ}, phase
field~\cite{DuLiuWang2004_JCP,BibenKassnerMisbah2005_PRE}, level
set~\cite{DoyeuxGuyotChabannesEtAl2013_JCAM}, boundary
integral~\cite{Shravan09,Rahimian15} and immersed boundary
approaches~\cite{KimLai2010_JCP,KimLai2012_PRE,HuLaiSeolEtAl2016_JCP}.
Most of these approaches assume a mathematical surface, whether
implicitly or explicitly, and define an elastic bending energy of the
surface. Researchers have developed a number of numerical methods for
calculating energy minimizing steady equilibrium shapes of lipid bilayer
membranes, vesicles and red blood cells. These approaches range from the
finite element method~\cite{Bartels,Peng13,RyKlYaCo16,Sinha15}, phase
field method~\cite{Du05,QiangDu08,Lowengrub13} to immersed boundary
method~\cite{Hu,Hu13, KimLai2010_JCP}. PI RR and collaborators led in
part the development of phase field functionals of membrane elastic
energy and approaches to coupling membrane elasticity to
fluids~\cite{0951-7715-18-3-016,Du05,DuEuler,QiangDu09}.

@@

Our HAP approach differs from prior methods in a number of respects.
First, we do not assume a surface. Rather, we more fundamentally
consider a collection of amphiphilic particles.  The collection of
amphiphiles minimize hydrophobic interactions by sequestering
hydrophobic tails in the form of a bilayer, and the particles' excess
free energy gives rise to an elastic bilayer energy. The second
difference lies in the fluid-interface coupling. Here, the associated
mobility problem~\eqref{eq:stokes} is more complicated than dealing with
a stress boundary condition or diffusive surface force, because the
fluid velocity must be that of a rigid body motion at the particle
surfaces. The HAP model also addresses the existence of multiple phases.
We can vary lipid length, spontaneous curvature, and bending rigidity by
introducing different particle shapes and hydrophobic boundary
conditions (Figure~\ref{fig:demixing}). In contrast, continuum theory
deals with multiple phases through additional surface densities that
must then satisfy specialized transport equations~\cite{Lowengrub07,
MikuckiZhou17}. 
%
\begin{wrapfigure}[14]{l}{0.35\textwidth}
\centerline{\includegraphics[width=0.35\textwidth]{figures/PW_fig2.pdf}}
  \vspace{-8pt}
  \caption{\label{fig:demixing} \footnotesize An initial assembly of
  small and large particles spontaneously segregates into two smaller
  bodies.}
\end{wrapfigure}
The greatest strength of using HAP to model lipid bilayer membrane is
the ability to form discontinuities (interfacial singularity) from first
physical principles without any artificial manipulation to rearrangement
the interface. For example, using the HAP model to simulate a vesicle
under a shear flow that is strong enough to cause the vesicle membrane
to rupture, we expect the HAP model to capture the reorganization of
lipid molecules on the scales of membrane thickness ($\sim 5$ nm), which
is a nearly impossible task using phase-field or immerse boundary method
without extreme refinement around the membrane and artificial treatment
of reconnection during the topological change of an interface.

>>>>

\begin{wrapfigure}[10]{r}{0.22\textwidth}
\centerline{\includegraphics[width=0.22\textwidth]{figures/BG_fig1.pdf}}
\caption{ \footnotesize
A collection $P_1,$ $P_2,$ $P_3,$ \dots of rigid particles.
The exterior domain $\Omega$ represents the solvent.}
  \label{fig:domain}
\end{wrapfigure}
Recently PIs RR and YNY developed a mathematical model called the
hydrophobic attraction potential (HAP) \cite{Fu2018_SIAM} that is based
on the physical origin of hydrophobicity.  This model addresses the
major shortcomings of molecular dynamics (MD) and continuum approaches.
Based on preliminary results (\S\ref{sec:preliminary_work}) the PIs
propose to extend this HAP model to offer an alternative modeling
methodology that leads to new mathematical ideas, and is both physically
accurate and computationally practical.
%
%
%The word hydrophobic (water fearing) derives from the low solubility of oil (hydrocarbon solute) in water and vice versa. 
%It causes hydrophobic moieties to aggregate and cluster,
%is responsible for the adhesion between hydrophobic surfaces \cite{Ducker2016}, large contact angles on a 
%dewetting surface \cite{Arenas2019,Sandre1999}, accumulation of particles along interfaces \cite{Lee2013,Lee2014}, 
%formation of micelles and bilayers \cite{Israelachvili80}, and protein folding and membrane insertion \cite{Kabelka2018}.
%
%Lipids are amphiphilic molecules whose  structure possesses
%both hydrophobic and hydrophilic parts. 
%The amphiphilic property is what allows lipids to form the membranes and 
%compartments of living cells \cite{Israelachvili80}. 
%More specifically, a lipid consists of 
%an elongated hydrocarbon tail that is hydrophobic, attached to a polar head that is hydrophilic.
%To shield the hydrophobic tails from water, lipids self-assemble into micelles and bilayers. 
%A micelle is a spherical arrangement of lipids with tails terminating at the micelle center. 
%A bilayer consists of two layers of lipids called monolayers, where the lipid tails point 
%from the monolayer surface into the bilayer core. 
%
%The mathematical modeling of a biological membrane is a challenging problem in applied mathematics. 
%Bilayers are elastic and resist deformations like bending, twisting, and stretching.
%Their elastic deformations are well described by the theory of liquid crystals \cite{ANDRIENKO2018520}.
%Lipid bilayer membranes can also be fluidic, and the lateral translation of lipids (or any membrane bound
%proteins) couples nonlocaly to the motion of the aqueous environment \cite{MerkelSackmannEvans1989,StoneAjdari1998_JFM,OppenheimerDiamant2009_BJ,OppenheimerDiamant2011_PRL}. 
%%Finally, the membranes of cellular 
%%compartments are constantly merging and pinching as part of intracellular trafficking. 
%%Therefore monolayer surfaces undergo discontinuous deformations. 
%
%There are two prevailing approaches in membrane modeling, and each has its advantages
%and disadvantages. Molecular dynamics (MD) is to date the only tool capable of resolving granular biological details
%The computational cost of MD, however, grows with the sixth power of the sample diameter and 
%so simulations are severely limited to 
%small system sizes and short time scales \cite{DiCarlo2019}.
%The other approach, continuum mechanics, assumes smooth surfaces and can therefore model 
%realistic systems over physical times. 
%But continuum description of a biological membrane ignores the granularity of lipid molecules, and thus requires some assumptions when a
%membrane ruptures or when two membranes fuse \cite{ChKo08}. 
%%
%%mechanics presumes, rather than predicts, the nucleation of discontinuities. 
%%
%
The proposed research aims to provide fundamental understanding of the self-assembly dynamics of amphiphilic particles. These results will facilitate optimal design of smart materials by
tuning the geometry and properties of the amphiphilic particles.

\subsection{Hydrophobic Attraction Potential (HAP)}
\label{sec:HAP}
%Based on the physical origin of hydrophobicity,  we have devised a
%functional, called the hydrophobic attraction potential (HAP), to model
%hydrophobic forces~\cite{Fu2018_SIAM}. 
The motivation for the HAP concept stemmed from PI RR's earlier work on
the problem of energy barriers in membrane
fusion~\cite{RyKlYaCo16,Chetal16}. By applying a squared gradient
mathematical theory for hydrophobic attraction between planar
surfaces~\cite{Eriksson1989,Lum1999,Menshikov2017,Marcelja1977}, PI RR
and collaborators resolved the long-standing issue of accounting for the
energy of a monolayer fissure surface during topological transitions.
Based on PI RR and YNY's coarse-grained membrane modeling
work~\cite{Fu2017}, the investigators initiated a collaboration to
create a gradient theory for arbitrary collections of hydrophobic and
amphiphilic particles. This new method eliminates the costly calculation
of water by treating the solvent implicitly, is capable of representing
biologically relevant morphologies, and avoids complicated
re-meshing schemes of continuum approaches by utilizing a particle-based
representation.

To define HAP, consider a collection of particles solvent suspended in a solvent. 
The region $\Omega \subset \mathbb{R}^3$  models 
the solvent phase (Figure \ref{fig:domain}).
%The HAP must be an energy of the shape of the region because hydrophobic
%attraction is non-additive.
The boundary of the region is the
water-particle interface.
Some parts of this interface are hydrophobic
while others are hydrophilic.
Hydrophobic interfaces disturb the hydrogen bond structure of water~\cite{Luzar1987,
Jonsson2006, Varilly2011}. This disturbance comes with an energetic penalty
that is proportional to interfacial area.
There is an additional decay length,
due to rapid fluctuation in the hydrogen bond
network, describing how far the restructuring extends into bulk water.

The above considerations motivate the following definition for the HAP:
\begin{align}
\label{HAP}
  \Phi = \gamma \int_{\Omega} \rho |\nabla u|^2 + \rho^{-1}u^2 \,\dif V. 
\end{align}
The integrand contains
the decay length $\rho$
and a scalar function $u(\mathbf{x})$ called the water activity.
The integrand has units of an inverse length and so
the volume integral has units of area. Multiplication by interfacial tension
$\gamma$ makes $\Phi$ an energy.

To get the effect of attraction, the water activity is not arbitrary, but rather
is the unique solution of the screened Laplace equation boundary value problem 
\begin{equation}
  \label{SL}
  -\rho^2 \Delta u + u = 0, \mbox{ } \mathbf{x} \in \Omega, \qquad
  u = f,  \mbox{ } \mathbf{x} \in \partial \Omega, \qquad 
  u(\mathbf{x}) \to 0 \mbox{ as } |\mathbf{x}| \to \infty.
\end{equation}
The boundary values $f$ define the degree of hydrophobicity of the
water-particle interface: $f=1$, describes a hydrophobic interface
and $f=0$ describes a hydrophilic interface. Thus $f$ encodes information
about the particles and the parameters $\rho$ and $\gamma$ encode information about the quality of the solvent 
~\cite{Israelachvili1954, Discher2002}. In practice, we use
\eqref{SL} to integrate by parts to obtain
\begin{equation}
\label{eq:HAP_easy}
\Phi = -\gamma \int_{\partial \Omega} \rho u \nabla u \cdot \boldsymbol{\nu} \,\dif S,
\end{equation}
thereby avoiding calculating a volumetric integral.

The equations~\eqref{HAP} and~\eqref{SL} possess number of the mathematical properties
that capture the phenomenological behavior of hydrophobic attraction.
Specifically, solutions of~\eqref{SL} yield an attractive interaction between
hydrophobic bodies separated by water ~\cite{Eriksson1989}.
Using boundary layer analysis, the HAP converges to a surface energy in the zero-decay length
limit~\cite{Lee2018, Lin2015, Shibata2004}.
Finally, we have demonstrated that the forces derived from HAP theory are non-additive~\cite{Meyer2006, Fu2018_SIAM}. 
%The hydrophobic force has been implicated in the directed folding of proteins, 
%adhesion between biological membranes, but it is still unanswered as to whether
%the hydrophobic force of the form (\ref{HAP}--\ref{SL}) exists between small molecules.

@@

To define the particle dynamics,
let $P_1$, $P_2$, \ldots, $P_N$,
be a finite collection of disjoint, rigid and closed particles
with Lipschitz boundary.
Assume that the label $f$ is an element of the Sobolev space $H^1(\Omega)$.
The functional~\eqref{HAP} has a minimizer among all functions $u$ equaling $f$ on the boundary in the sense of trace.
Conversely, from maximum principles and energy estimates, the solution of~\eqref{SL} is unique and minimizes~\eqref{HAP}.
Taking the first variation of~\eqref{HAP}, i.e. the derivative with respect to the domain
~\cite{Bandle2015, Schiffer1954, Grinfeld2010},
yields a symmetric, rank-two tensor called the hydrophobic stress:

@@

  \mathbf{F}_{\text{hydro},i} = \iint_{\partial P_i} \boldsymbol{\sigma}_{\text{hydro}}
  \cdot \boldsymbol{\nu} \,\dif S,\quad
  \mathbf{G}_{\text{hydro},i} = \iint_{\partial P_i} (\mathbf{x}-\mathbf{a}_i) \times
  (\boldsymbol{\sigma}_{\text{hydro}} \cdot \boldsymbol{\nu}) \,\dif S

@@

relative to the center of mass $\mathbf{a}_i$.
This system is force- and torque-free  \S\ref{subsec:specific_aim_2}.
To avoid particle collisions, we define an excluded volume potential $\Phi_{\text{repul}}$
that diverges whenever tubular neighbordhoods of adjacent particles overlap,
along with the excluded volume repulsions. The total potential, force and torque are then

@@

\label{eq:total_poten}
\Phi = \Phi_{\text{hydro}} + \Phi_{\text{repul}},\quad
\mathbf{F}_{i} = \mathbf{F}_{\text{hydro},i} + \mathbf{F}_{\text{repul},i},\quad
\mathbf{G}_{i} = \mathbf{G}_{\text{hydro},i} + \mathbf{G}_{\text{repul},i}.

@@

To supply viscous dissipation, we incorporate the mobility problem in
Stokes flow for a rigid body suspension:
\begin{equation}
\label{eq:stokes}
\begin{aligned}
  &-\mu \Delta \mathbf{u} + \nabla p = 0, \quad \mathbf{x} \in \Omega, \qquad 
  \nabla \cdot \mathbf{u} = 0,  \quad \mathbf{x} \in \Omega,\\
  &{\bf u}(\mathbf{x}) \to 0 \quad \text{as}\ |\mathbf{x}|\to \infty,\qquad 
  \mathbf{u}(\mathbf{x})|_{\partial P_i} = \mathbf{v}_i +
\boldsymbol{\omega}_i\times(\mathbf{x} - \mathbf{a}_i),\\
&\int_{\partial P_i}\boldsymbol{\sigma}\cdot {\bf n} \dif S=-{\bf F}_i, \quad
\int_{\partial P_i} (\mathbf{x} - \mathbf{a}_i)\times (\boldsymbol{\sigma} \cdot \mathbf{n}) \dif S =-{\bf G}_i.
\end{aligned}
\end{equation}
Here $\mu$ is the fluid viscosity and the first two equations state that
the fluid motion is a divergence-free Stokes flow; the third equation
specifies that the fluid velocity vanishes at infinity; the fourth
equation enforces a rigid body motion on each particle, where
$\mathbf{v}_i$ and $\boldsymbol{\omega}_i$ are unknown translation and
angular velocities; the last two equations state that the net
viscous forces and torques with fluid shear stress $\boldsymbol{\sigma}$
balance \eqref{eq:total_poten}.

The time integration of
particle configurations goes as follows: \textbf{(i)} solve the
BVP~\eqref{SL} for the screened Laplace equation, \textbf{(ii)}
determine the rigid body forces and torques~\eqref{forceandtorque},
\textbf{(iii)} solve the Stokes mobility problem~\eqref{eq:stokes} for
rigid body motion, \textbf{(iv)} update the particle configuration.
Numerical challenges and how these challenges are overcome are described
in \S\ref{subsec:specific_aim_2}.

@@

The HAP formulation is, to our knowledge, the first demonstration of bilayer self-assembly
by a continuum-based interaction model~\cite{Noguchi2001, Farago2003, Brannigan2006,
Brooks2009, Wang2013}.
Our simulations use Janus particles to model lipid amphiphiles which are
popular in material science and physics for creating functional
materials~\cite{Lee2014, Lee2013}. Janus particles are typically
spherical with a biphasic material label on either hemisphere, endowing
the particle with a directional order.
We model elongated lipid by elliptical particles with the hydrophobic label defined  along the ellipse's axis.
Under the hydrophobic force, and with an excluded volume,
the Janus particles spontaneously merge and realign to form
bilayers. This occurs only as a result of energy minimization and does
not require artificial inputs.

It is worth emphasizing that the HAP model uses only a few parameters;
interfacial tension, decay length, repulsion strength, and particle
shape. For example, an elastic modulus for stretching a vesicle from
micropipette manipulation directly calibrates our interfacial tension
parameter. This is in direct contrast with pair-potential based
approaches in MD simulations and coarse-grained models where many more
parameters are required~\cite{Varilly2011, Wang2013}.
%MD simulators have also made measurements and lately there is better and better agreement with reality. 
%But even the simplest coarse grained models based on pair potentials for lipids has many more parameters \cite{Varilly2011,Wang2013} .

As a proof of concept, our work has already tested for elastic energies
for bending, stretching, and tilt of the bilayer assembly. 
The elastic coefficients derived from the HAP simulations show strikingly positive
agreement with experimentally determined values~\cite{Fu2018_SIAM}.
Encouraged by these results and the hydrodynamic simulations of the
bilayer assembly, the PIs propose to extend the HAP model
to make direct comparison with experimental results  \S\ref{subsec:specific_aim_1},
to establish that the HAP model has the capability
and adaptivity to model flow phenomena across length scales and time scales \S\ref{subsec:specific_aim_2},
using efficient, high-order numerical methods \S\ref{sec:proposed-work}.

\begin{wrapfigure}[14]{l}{0.35\textwidth}
\centerline{\includegraphics[width=0.35\textwidth]{figures/PW_fig2.pdf}}
  \vspace{-8pt}
  \caption{\label{fig:demixing} \footnotesize An initial assembly of
  small and large particles spontaneously segregates into two smaller
  bodies.}
\end{wrapfigure}Over the past decades, researchers have used a number of mathematical
tools to simulate vesicles in a shear flow, including lattice
Boltzmann~\cite{KaouiHartingMisbah2011_PRE}, coarse-grained Brownian
dynamics~\cite{NoguchiTakasu2002_BJ}, phase
field~\cite{DuLiuWang2004_JCP,BibenKassnerMisbah2005_PRE}, level
set~\cite{DoyeuxGuyotChabannesEtAl2013_JCAM}, boundary
integral~\cite{Shravan09,Rahimian15} and immersed boundary
approaches~\cite{KimLai2010_JCP,KimLai2012_PRE,HuLaiSeolEtAl2016_JCP}.
Most of these approaches assume a mathematical surface, whether
implicitly or explicitly, and define an elastic bending energy of the
surface. Researchers have developed a number of numerical methods for
calculating energy minimizing steady equilibrium shapes of lipid bilayer
membranes, vesicles and red blood cells. These approaches range from the
finite element method~\cite{Bartels,Peng13,RyKlYaCo16,Sinha15}, phase
field method~\cite{Du05,QiangDu08,Lowengrub13} to immersed boundary
method~\cite{Hu,Hu13, KimLai2010_JCP}. PI RR and collaborators led in
part the development of phase field functionals of membrane elastic
energy and approaches to coupling membrane elasticity to
fluids~\cite{0951-7715-18-3-016,Du05,DuEuler,QiangDu09}.

@@

Our HAP approach differs from these prior methods in a number of respects.
First, we do not assume a surface. Rather, we more fundamentally
consider a collection of amphiphilic particles.  The collection of
amphiphiles minimize hydrophobic interactions by sequestering
hydrophobic tails in the form of a bilayer, and the particles' excess
free energy gives rise to an elastic bilayer energy. The second
difference lies in the fluid-interface coupling. Here, the associated
mobility problem~\eqref{eq:stokes} is more complicated than dealing with
a stress boundary condition or diffusive surface force, because the
fluid velocity must be that of a rigid body motion at the particle
surfaces. The HAP model also addresses the existence of multiple phases.
We can vary lipid length, spontaneous curvature, and bending rigidity by
introducing different particle shapes and hydrophobic boundary
conditions (Figure~\ref{fig:demixing}). In contrast, continuum theory
deals with multiple phases through additional surface densities that
must then satisfy specialized transport equations~\cite{Lowengrub07,
MikuckiZhou17}.

The greatest strength of using HAP to model lipid bilayer membrane is
the ability to form discontinuities (interfacial singularity) from first
physical principles without any artificial manipulation to rearrangement
the interface. For example, using the HAP model to simulate a vesicle
under a shear flow that is strong enough to cause the vesicle membrane
to rupture, we expect the HAP model to capture the reorganization of
lipid molecules on the scales of membrane thickness ($\sim 5$ nm), which
is a nearly impossible task using phase-field or immerse boundary method
without extreme refinement around the membrane and artificial treatment
of reconnection during the topological change of an interface.
